---
title: "Outlier Identification"
vignette: >
  %\VignetteIndexEntry{0A_Outlier Identification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  knitr:::html_vignette:
    toc: yes  
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  digits = 3,
  collapse = TRUE,
  comment = "#>"
  )
options(digits = 3)
```


This package implements the method by Newman, D. The test statistic q=w/s where w is the range of the data and s is sample standard deviation estimated from controls. The null distribution for q has been derived and 1% and 5% quantiles have been given in the paper.

We implement this procedure. In particular, `NewmanTest()` returns a logic vector specifying whether the observations are outliers.

Usually, drug-response data contains multiple doses. Therefore, we write a wrapper `drOutlier()` that compute the result for all doses one dose at a time.

We use the *ryegrass* data from **drc** package for illustration purpose. The *ryegrass* data was originally published by [(Streibig et al. 2002)](#Streibig2002ryegrass) which contains 24 concentrations of ferulic acid (a root growth inhibitor) and corresponding root length. 

First, we load the **drexplorer** package and attach the \Rpackage{ryegrass} data.

```{r, message = FALSE}
library(drexplorer2)

data(ryegrass)
dose <- ryegrass[, 2]
response <- ryegrass[, 1]
```

At dose=3.75 and significance level 0.05, we find there is one outlier identified:

```{r}
## potential outlier at dose 3.75
NewmanTest(ref = response[dose == 0], obs = response[dose == 3.75], alpha  =0.05)
```

We also examine all dose levels and find no further outliers:

```{r}
drOutlier(drMat = ryegrass[, c(2, 1)], alpha = 0.05)
```
