---
title: "Assessing Reproducibility"
vignette: >
  %\VignetteIndexEntry{0B_Assessing Reproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  knitr:::html_vignette:
    toc: yes
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  digits = 3,
  collapse = TRUE,
  comment = "#>"
  )
options(digits = 3)
```

Sometimes replicated viability assays are performed. In such case, it is useful to examine if the experiments are reproducible.

A good metric is the **Concordance Correlation Coefficient (CCC)** that captures both the location shift as well as scale shift between the replicates. The `plotCCC` function can be used to compute CCC and visualize the replicated data.
 

```{r, fig.height=5, fig.width=5, dev='png', dpi=96*10, fig.align='center'}
library(drexplorer2)

set.seed(100)
r1 <- runif(28)
r2 <- r1+rnorm(28, 0, 0.1)
ccc <- plotCCC(r1, r2,
               xlab = 'Simulated response, replicate 1',
               ylab = 'Simulated response, replicate 2')
ccc
``` 

Here we have simulated two response vectors and calculate CCC. The computed CCC value is `r round(ccc['ccc'], 3)`, location shift is `r round(ccc['l_shift'], 3)`, scale shift is `r round(ccc['s_shift'], 3)`, Pearson correlation is `r round(ccc['corr'], 3)`. 
