---
title: "Select the Best Model"
vignette: >
  %\VignetteIndexEntry{Select the Best Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  knitr:::html_vignette:
    toc: yes  
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  digits = 3,
  collapse = TRUE,
  comment = "#>"
  )
options(digits = 3)
```

We provide a wrapper function, `fitOneExp()`, incorporating outlier detection, fitting multiple models, selecting the best model and estimating IC values. 

Users can specify multiple models from the **drc** and **DoseFinding** packages using the *models* parameter. An example is shown below.

```{r, eval=TRUE}
library(drexplorer2)
data(ryegrass)

fitExp <- fitOneExp(ryegrass[, c(2, 1)],
                    drug = '', cellLine = '', unit = '',
                    models=c('sigEmax', 'LL.4', 'LL.5', 'LL.3', 'LL.3u', 'logistic'),
                    alpha = 0.05, interpolation = TRUE)
```

The fitted models can be compared through the plotOneExp function as below:

```{r, warning=FALSE, fig.height=6, fig.width=6, dev='png', dpi=72*10, fig.align='center'}
plotOneExp(fitExp, main = 'Dose-Response Curves from Multiple Models')

```

The best model based on Residual Standard Error (RSE) can be obtained as shown below.

```{r}
## This is the bast model:
bM <- fitExp$bestModel
bM

## RSE from all models fitted:
fitExp$RSEs

## IC50 from all models fitted:
fitExp$IC50

## best model fitted:
fitExp$fit[[bM]]@fit
```




